\section{Approximation Algorithms: Disposition}
\begin{enumerate}
	\item Performance ratio
	\item Vertex-cover
	\item Traveling-salesman
	\item Set-covering
	\item Randomization \& Linear programming
	\item Subset-sum problem
\end{enumerate}

CLRS 35.1, 35.2, 35.3, 35.4, 35.5

\section{Approximation Algorithms: Notes}
An approximation algorithm is an algorithm that computes a near-optimal solution,
often because the optimal solution is NP-complete and thus intractable or at least
exponential in running time. One solution to that is to only work on small problem sets,
where exponential time is fine. Another is to compute a near-optimal solution using an
approximation algorithm.

\subsection{Performance ratio}
We say that an algorithm for a problem has an \textbf{approximation ratio} of
$p(n)$ if, for any input of size $n$, the cost $C$ of the solution produced by
the algorithm is within a factor of $p(n)$ of the cost $C^*$ of an optimal
solution:
%
\[
	max\left( \frac{C}{C^*}, \frac{C^*}{C}\right) \leq p(n)
\]
%
If an algorithm achieves an approximation ratio of $p(n)$ we call it a
$p(n)$\textbf{-approximation algorithm}. Because we assume all solutions have
a positive cost, these ratios are always well-defined. The approximation ratio of
an approximation algorithm is never less than 1, since $C/C^*\leq 1$ implies $C^*/C \geq 1$.
Therefore, a 1-approximation algorithm produces an optimal solution, and an approximation
algorithm with a large approximation ratio may return a solution much worse than optimal.

Below some schemes are outlined which allow us to give a value $\epsilon$ along
with the instance of the problem and achieve an approximation that have a
quality depending on the value.
\begin{description}
\item[Approximation Scheme] for an optimization problem is an approximation
  algorithm that takes as input, not only an instance of the problem, but also a
  value $\epsilon > 0$ such that for any fied $\epsilon$ the scheme is a
  $(1+\epsilon)$-approximation algorithm.

\item[Polynomial-Time Approximation Scheme] is an approximation scheme if for
  any fixed $\epsilon > 0$, the scheme runs in polynomia ltime in the size $n$
  of it's input instance. The running time of such a scheme can increase rapidly
  as $\epsilon$ decreases. For example, the running time of a polynomial-time
  approximation scheme might be $O(n^{(2/\epsilon)})$. Ideally, if $\epsilon$
  decrease by a constant factor, the running time to achieve the desired
  approximation should not incrase by more than a constant factor. (Not
  necessarily by the same factor $\epsilon$ was decreased with.)

\item[Fully Polynomial-Time Approximation Scheme] means the approximation
  algorithm runs in polynomial time in both $1/\epsilon$ and the size $n$ of the
  input instance. I.e. $O((1/\epsilon)^2n^3)$. With such a scheme, a constant
  factor decrease in $\epsilon$ comes with a constant factor increase in
  runningtime.
\end{description}

\subsection{Vertex-cover}
A Vertex Cover of an undirected graph $G = (V,E)$ is a subset $V' \subseteq V$
such that if $(u,v)$is an edge of $G$, then either $u \in V'$ or $v \in V'$ or
both. The size of a vertex cover is the number of vertices in it.

The Vertex Cover Problem is to find a vertex cover of minimum size in a given
undirected graph, this is an optimization version of an NP-Complete decision
problem.

\begin{description}
\item[Theorem 35.1] \texttt{Approx-Vertex-Cover} is a polynomial-time
  2-approximation algorithm.
\item[Proof] It is already shown that the algorithm is a polynomial time
  algorithm.

  The set $C$ of vertices return must be a vertex cover since it loops until
  every edge in $G.E$ have been covered by some vertex.

  To see that the algorithm returns a vertex cover of at most twice the size of
  an optimal cover, let $A$ denote the set og edges that line 4 selects.  To
  cover all the edges in $A$, any vertex cover, in particular the optimal vertex
  cover $C^*$, must include at least one endpoint of each edge in $A$. No two
  edges share endpoints since when an edge is picked, all edges incident on its
  endpoints are removed. Thus no two edges in $A$ are covered by the same vertex
  from $C^*$, and we have the lower bound
  \[
    |C^*| \geq |A|
  \]
  on the size of an optimal vertex cover. Each execution of line 4 picks an edge
  which neither of its endpoints are already in $C$, yielding an upper boind on
  the size of the vertex cover returned
  \[
    |C| = 2|A|
  \]
  Combining the two bounds gives
  \begin{align*}
    |C| &= 2|A| \\
        &\leq 2|C*|
  \end{align*}
  thereby proving the theorem. \qed
\end{description}

\subsection{Travelling-salesman}
The TSP is also NP-Complete, the following is a method for approximating an
optimal solution that is at most twice as long as the optimal tour.

It creates a minimum spanning tree using Prims algorithm, and walks along this
tree using a pre-order walk (parent first then child), and uses this as the
solution.

\begin{description}
\item[Theorem 35.2] \texttt{Approx-TSP-Tour} is a polynomial-time
  2-approximation algorithm for the traveling salesman problem, with the
  triangle inequality.

\item[Proof] See page 1114.
\end{description}

\subsection{Set-covering}
\subsection{Randomization \& Linear programming}
\subsection{Subset-sum problem}