\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[12pt]{article}
\usepackage[utf8]{inputenc} 
\usepackage[T1]{fontenc}
\usepackage[english]{babel} 
\usepackage[margin=2.5cm]{geometry}
\geometry{a4paper}
\usepackage{longtable}


%----------------------------KODE START---------------------------------------
\usepackage{listings}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\lstset{
  basicstyle=\footnotesize,
  numbers=left,
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  numbersep=10pt,
  backgroundcolor=\color{white},
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,                   % adds a frame around the code
  rulecolor=\color{black},        
  tabsize=4,
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  title=\lstname,                   % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{mauve},          % keyword style
  commentstyle=\color{Maroon},       % comment style
  stringstyle=\color{BlueViolet},         % string literal style
  escapeinside={\%*}{*)},            % if you want to add LaTeX within your code
  morekeywords={*,...},              % if you want to add more keywords to the set
  deletekeywords={...}              % if you want to delete keywords from the given language
}
%----------------------------KODE SLUT----------------------------------------


\setlength\parindent{0pt} % Makes \noindent standard
\usepackage{graphicx} 
\usepackage{sidecap}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float} 
\usepackage{wrapfig} % Allows in-line images if needed
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\hypersetup{colorlinks=false,hidelinks, citecolor=black, urlcolor=black}
\usepackage{csquotes}
\usepackage{comment}

\usepackage[dot, autosize, outputdir="dotgraphs/"]{dot2texi}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usepackage{url}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{longtable}
\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}
\usepackage[titletoc]{appendix} % Names appendices "Appendix A"
                                % instead of just A in Contents
\usepackage[bottom]{footmisc}
\usepackage{pdfpages}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx


%\usepackage{lmodern}
\usetikzlibrary{arrows,automata}
\usepackage{verbatim}

\linespread{1.2} 
\graphicspath{{./figures/}} 

% fancy drawings
\usepackage{pgf}
\usepackage{epigraph}

% \epigraphsize{\small}% Default
\setlength\epigraphwidth{8cm}
\setlength\epigraphrule{0pt}

\usepackage{etoolbox}

\makeatletter
\patchcmd{\epigraph}{\@epitext{#1}}{\itshape\@epitext{#1}}{}{}
\makeatother

%\usepackage{boxproof}
%\usepackage{nomencl}
\usepackage{natbib}

\newcommand{\fasto}{\textsc{Fasto} }
\newcommand{\mips}{\textsc{Mips} }
\newcommand{\mars}{Mars }
\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

%-----------------------------------------------------------------------------
\begin{document}
\begin{titlepage}

\centering 

\textsc{\Large Advanced Algorithms and Datastructures}\\[0.5cm] 
\textsc{\large Exam Notes}\\[0.5cm] 

\vfill

\emph{Author:}
\\
Jenny-Margrethe \textsc{Vej} -- 250986 -- rwj935
\vspace{20mm}

{\large Block 4, April-June, 2014}\\[3cm] 

\end{titlepage}
%-----------------------------------------------------------------------------

\tableofcontents
\newpage

%-----------------------------------------------------------------------------

%%Completely done 
\input{MaxFlow}
\newpage

%%Missing the last 4 pages in the chapter in the book and the disposition
\input{FibHeaps}
\newpage

%%
\input{NP-Complete}
\newpage

\input{Randomised}
\newpage


%%Haven't looked at those below here

\section{Hashing}
OMG LOLZ'n SHIZZLES! You'r gonna nail this shit!

\subsection{Direct-address tables}
We have a scenario where we want to maintain a dynamic set of some sort. Each element has a key drawn from a universe $U = \{0, 1, ..., m-1\}$ where $m$ is not too large, and no two elements have the same key. \\

Representing it by a direct-address table, or array $T[0...m-1]$, each slot or position corresponds to a key  in $U$. If there is an element $x$ with the key $k$, then $T[k]$ contains a pointer to $x$ - otherwise $T[k]$ is empty represented by $NIL$\\

Illustration in the book on page 254
\begin{algorithm}
\begin{algorithmic}[1]
\State{DIRECT-ADDRESS-SEARCH}{$(T, k)$}
   \State \textbf{return} $T[k]$\\
   
\State{DIRECT-ADDRESS-INSERT}{$T, x$}
   \State \textbf{return} $T[key[x]] \leftarrow x$\\

\State{DIRECT-ADDRESS-DELETE}{$T, x$}
   \State \textbf{return} $T[key[x]] \leftarrow NIL$
\end{algorithmic}
\end{algorithm}

Downsides of direct addressing: If the universe $U$ is large, storing a table $T$ of size $|U|$ may be impractical or even impossible, given the memory available on a typical computer. Furthermore, the set $K$ of keys actually stored may be so small relative to $U$ that most of the space allocated for $T$ would be wasted. 

\subsection{Hash tables}
To overcome the downsides of direct addressing, we can use hash tables. When $K$ is much smaller than $U$, a hash table requires much less space than a direct-address table. It can also reduce storage requirements to $\Phi (|K|)$. \\

This bound is for average-case time whereas for direct addressing it holds for the worst-case time. \\

With direct addressing, an element with $k$ is stored in slot $k$. With hashing, this element is stores in slot $h(k)$; that is, we use a hash function $h$ to compute the slot from the key $k$. There is one hitch: 2 keys may hash to the same slot - called a collision. See page 256 for illustrations. \\

A hash function $h$ must be deterministic in that a given input $k$ should always produce the same output $h(k)$. \\

\textbf{Collision resolution by chaining}\\
In chaining we place all the elements that hash to the same slot into the same linked list. See figure on page 257.

\begin{algorithm}
\begin{algorithmic}[1]
\State{DIRECT-HASH-SEARCH}{$(T, x)$}
   \State \textbf{search} for an element with key $k$ in list $T(T, k)$\\
   
\State{DIRECT-HASH-INSERT}{$T, x$}
   \State \textbf{insert} $x$ at the head of the list $T[h(x.key)]$\\

\State{DIRECT-HASH-DELETE}{$T, x$}
   \State \textbf{delete} $x$ from the list $T[h(x.key)]$
\end{algorithmic}
\end{algorithm}

\textbf{Analysis of hashing with chaining}\\
Given a hash table $T$ with $m$ slots that stores $n$ elements, we define the load factor $\alpha$ for $T$ as $n/m$ which is the same as the average number of elements stores in a chain. \\

Worst case performing: $\Phi (n)$ + the time to compute the hash function (all $n$ keys hash to the same slot, creating a list of length $n$. \\

Average case performing of hashing depends on how well the hash function $h$ distributes the set of keys to be stores among the $m$ slots, on the average. \\

\textbf{\textit{Theorem 11.1}}
In a hash table in which collisions are resolved by chaining, an unsuccessful search takes an average-case time $\Phi (1 + \alpha)$, under the assumption of simple uniform hashing. (OBS: Beviset herfor står på side 259!! LÆS OG FORSTÅ)\\

\textbf{\textit{Theorem 11.2}} 
In a hash table in which collisions are resolved by chaining, a successful search takes average-case time $\Phi (1 + \alpha)$, under the assumption of simple uniform hashing. (OBS: Beviset herfor står på side 259-260!! LÆS OG FORSTÅ)

\subsection{Hash functions}
\textbf{What makes a good hash function?} \\
A good hash function satisfies (approximately) the assumption of simple uniform hashing: each key is equally likely to hash to any of the $m$ slots, independently of where any other key has hashed to. In practice, we can often employ heuristic techniques to create a hash function that performs well. \\

\textbf{Interpreting keys as natural numbers}\\
Most hash functions assume that the universe of keys is the set $\mathbb{N} = \{0, 1, 2, ...\}$ of natural numbers. 

\subsubsection{The division method}
\subsubsection{The multiplication method}
\subsubsection{Universal hashing}
\subsection{Open addressing}
\subsection{Perfect hashing}
\newpage

\section{Exact Exponential Algorithms}
\newpage

\section{Approximation Algorithms}
\newpage

\section{Computational Geometry}
\newpage

\section{Linear Programming and Optimisation}
[\cite{monoids}]

\newpage
\bibliographystyle{plainnat}
%\bibliographystyle{unsrt}
\bibliography{bibliography}


%-----------------------------------------------------------------------------
\end{document}